---
title: "How does a Decision Tree work?"
format: 
    revealjs:
        embed-resources: true
editor: visual
author: René Siodlaczek
date: February 17, 2025
---

## About

### Goal

The app aims to support learning how the decision tree classification algorithm works.

### How it works

-   In the app the user can train a decision tree classifier with different hyperparameters.

-   The user sees a visual representation of the fitted decision tree.

-   The decision tree can be used to make predictions.

## Data

The underlying dataset is called 'Palmer Archipelago (Antarctica) Penguin Data'. It contains features from different penguins observed in the Palmer Archipelago near Palmer Station, Antarctica.

```{r}
library(palmerpenguins)
library(DT)

DT::datatable(penguins)
```

## Training

After the users selected the algorithm's hyperparameter and clicked on "Train decision tree" the algorithms tries to find the best decision tree to predict a penguin's species.

Once the training is complete a visual decision tree appears:

```{r}
library(rpart)
library(rpart.plot)

penguins$year <- as.factor(penguins$year)

model <- rpart::rpart(species ~ .,
                     data = penguins,
                     na.action = na.rpart,
                     method = "class",
                     parms = list(split = "gini"),
                     minsplit = 10,
                     cp = 0.01,
                     maxdepth = 6)

rpart.plot::rpart.plot(model)
```

## Prediction

The trained model can then be used to predict the species of a new penguin.

```{r}
#| echo: true
newdata <- data.frame(
    island = "Biscoe",
    bill_length_mm = 43,
    bill_depth_mm = 17,
    flipper_length_mm = 200,
    body_mass_g = 4201,
    sex = "female",
    year = "2007")

predict(model, newdata = newdata)
```
